---
title: "STAT 2450 Assignment 3 (40 points) "
author: "Emma Leitao"
date: 'Banner:  B00713241'
output:
  word_document: default
  html_document: default
  pdf_document: default
---


1.  Write a function to calculate miles per gallon given kilometres travelled, and litres of gasonline used.  The function should have two arguments, litres and kilometres, and should return the mileage in mpg.

```{r}
mpg=function(litres,kilometers){
  miles = kilometers*5/8
  gallons = litres/4.55
  mpg=miles/gallons
  return(mpg)
}

mpg(8.3,100)
```

Test your function using input values of 100 kilometres and 8.3 litres.


(5 points)


2.  The roots of the quadratic $a x^2 + b x + c$ are given by

$$\frac{-b \pm \sqrt{b^2-4ac}}{2a}$$

If $b^2-4ac <0$, the quadratic has no real roots.

Write a function to calculate the real roots of a quadratic.  The function
should have 3 arguments, *a*, *b* and *c*.
If $b^2-4ac <0$, the function should print "quadratic has no real roots",
and then return(NULL).
Otherwise, the function should return a vector of length 2, those being
the real roots (which may be the same if  $b^2-4ac =0$).

Test your function using the quadratic $x ^2 - 3 x + 2$.
```{r}
roots = function(a,b,c){
  disc = b^2-(4*a*c)
  if(disc >= 0){
    roots = (-b + c(sqrt(disc),-sqrt(disc)))/2*a
  }else{
    roots = "There are no real roots"
  } 
  return(roots)
} 
roots(1,-3,2)
```


(5 points)

3.  Where $x_1, x_2, \ldots , x_n$ is a sample from a normal distribution
with unknown mean $\mu$ and unknown variance $\sigma^2$, the level
$100(1-\alpha)$% confidence interval for $\mu$ is given by

$$\bar x \pm t_{1-\alpha/2, n-1} \frac{s}{\sqrt n}$$

where $\bar x$ and $s$ are the sample mean and sample standard deviation
of the data, and $t_{1-\alpha/2, n-1}$ cuts off an area $1-\alpha/2$ to
its left under the $t$ curve with $n-1$ degrees of freedom.

Write a function which has two arguments, a vector of data $x$, and
alpha, which takes any value between 0 and 1, but should have a default value of .05 (hence you must learn how to program functions with default values for their arguments).

The function should return a vector of length 2, which contains the endpoints of the confidence interval.


```{r}
cint <- function(x,alpha = 0.05){
  if(alpha >1 | alpha < 0){
    return("alpha should be between 0 and 1")
    break
  }
  xbar <- mean(x, na.rm = T)
  xsd <- sd(x, na.rm = T)
  n <- length(x)
  qtlower <- qt(alpha/2,n-1)
  qtupper <- qt(1-alpha/2,n-1)
  cint <- xbar+c(qtlower,qtupper)*(xsd/sqrt(n))
  return(cint)
}

```

The percentiles of the t-distrubtion can be calculated as follows. Suppose
that you want the 97.5'th percentile of the t-distribution with 23 degrees
of freedom.  This can be calculated in R as

```{r}
qt(.975,23)
```

Test your function by calculating the 99% confidence interval using
the following data

```{r}
set.seed(87612345)
data=rnorm(25,mean=4.5,sd=.75)
```
You can check your calculation using

```{r}
t.test(data,conf.level=0.99)
cint(data,0.01)
```

When putting your two endpoints together, you may find something similar to the following to be useful.

```{r}
1+c(-1,1)*.25
```

(5 points)

4.
The derivative of a function $f(x)$ at $x$ can be approximated by the Newton's quotient

$$\frac{f(x+h) - f(x)}{h}$$

where $h$ is a small number.  Write a function to calculate the Newton's quotient for the function $f(x) = exp(x)$.  The function should take two scalar arguments, $x$ and $h$.
Use a default value of $h=1.e-6$.
Test your function at the point $x=1$ using the default value of $h$, and compare to the true value of the derivative $f'(1) = e^1$.

```{r}
nquot <- function(x,h = 1.e-6){
  nquot <- (exp(x+h)-exp(x))/h
  return(nquot)
}

nquot(1)
exp(1)
```


(5 points)

5.  A very useful feature in R is the ability to pass a function name as an argument.
Here is an example, where 2 is added to the value of a function, for three different functions $exp(x)$, $log(x)$, and $sin(x)$, at selected points $x$.

```{r}
test=function(x,f){
 output=f(x)+2
 return(output)
}

test(0,exp)
test(1,log)
test(0,sin)
test(pi/2,sin)
```

Modify your function from problem 4 so that you pass in the name of the function for which you want to approximate the derivative.  Use the same default value for $h$, and approximate
the derivative of sin(x) at $x=\pi/4$, of $log(x)$ at $x=2$, and of $exp(x)$ at $x=1$.

```{r}
nquot <- function(f,x,h = 1.e-6){
  nquot <- (f(x+h)-f(x))/h
  return(nquot)
}

nquot(sin,pi/4)
nquot(log,2)
nquot(exp,1)

```

(10 points)


6. 
Write a function which takes one argument $x$ of length 2, and returns the ordered values of $x$.  That is, if $x_2<x_1$, your function should return $c(x_2,x_1)$, otherwise it should return $x$.  (WRITE YOUR OWN FUNCTION. DO NOT USE THE BUILT IN FUNCTION ORDER)

Use your function to process a dataset with 2 columns as follows.  Iterate over the rows of the data set, and if the element in the 2nd column of row *i* is less than the element in the first
column of row *i*, switch the order of the two entries in the row by making a suitable call to the function you just wrote.

Test using the following data.

```{r}
set.seed(1128719)
data=matrix(rnorm(20),byrow=T,ncol=2)

data
```

```{r}
orderx <- function(x){
  for(i in 1:nrow(x)){
    if(x[i,2]<x[i,1]){
      x[i,] <- c(x[i,2],x[i,1])
    }
  }
  return(x)
}

orderx(data)
```

(10 points)